---
title: "Pset 4"
author: "Igor Morzan"
date: "2/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# these are the necessary libraries with their functions for this PSET
library(gt)
library(tidyverse)
library(janitor)
library(lubridate)

data <- read_csv(file = "ps_4_elections-poll-nc09-3.csv",
  col_types = cols(
    .default = col_character(),
    turnout_scale = col_double(),
    turnout_score = col_double(),
    w_LV = col_double(),
    w_RV = col_double(),
    final_weight = col_double(),
    timestamp = col_datetime(format = "")))

data %>%
  clean_names() %>%
  filter(!is.na(response), !is.na(race_eth), !is.na(final_weight), educ != "[DO NOT READ] Refused")
```

## Question 1
```{r}
madlib_1 <- data %>% 
  filter(response == "Dem") %>%
  count()
```
There were `r madlib_1` respondents who supported the Democratic candidate.

```{r}
madlib_2_dataframe <- data %>%
  filter(response %in% c("Rep", "Und")) %>%
  group_by(response) %>%
  count()

madlib_2 <- madlib_2_dataframe$n[1] - madlib_2_dataframe$n[2]
```
There were `r madlib_2` more respondents who favored the Republican candidate than who were Undecided.

```{r}
madlib_3 <- data %>%
  select(gender, gender_combined) %>%
  filter(gender != gender_combined) %>%
  count()
```
There are two gender variables (gender and gender_combined). There are `r madlib_3` individuals for whom these variables have different values.

```{r}
madlib_4 <- data %>%
  filter(race_eth == "White") %>%
  filter(file_race_black != "White") %>%
  count()
```
There are `r madlib_4` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

```{r}
madlib_5 <- data %>%
  filter()
```
The first response of Rep came X minutes (rounded to the nearest minute) before the first response of Dem.

## Question 2
```{r make_chart, echo=FALSE, results="asis"}

# You need results="asis" anytime you want to use gt to create a chart.

data %>% 
  select(response, educ4, final_weight) %>% 
  group_by(educ4, response) %>% 
  
  # Again, this is not a course in survey weighting. There is an argument that I
  # should just ignore the topic altogether. But, I really like to replicate
  # published results, and that always requires weights. Hence today's
  # monologue. But only two people fell asleep during it --- not kidding! --- so
  # I count that as a victory.
  
  # All you need to know for this class is: Use sum(weight_var) in place of n().
  
  summarize(total = sum(final_weight)) %>%   
  filter(educ4 != "[DO NOT READ] Don't know/Refused") %>% 
  spread(key =  response, value = total) %>% 
  mutate(all = Dem + Rep + Und) %>% 
  mutate(Dem = Dem / all) %>% 
  mutate(Rep = Rep / all) %>% 
  mutate(Und = Und / all) %>% 
  select(-all) %>% 
  
  # One of the biggest pieces of black magic incantation in R is ungroup(). (I
  # did not mention this in class.) Summary: Whenever you group a tibble (as we
  # do above) the grouping stays with an resulting object, until you explicitly
  # ungroup() it. That can't ever hurt things (right? TFs?) and it often helps,
  # as in this case.
  
  ungroup() %>% 
  
  # You will have a chance to explore many other gt commands in problem set #4.
  # I added two extras that we did not get to in class.
  
  gt() %>% 
    tab_header(
      title = "Polling Results in Arizon 6th Congressional District") %>% 
    
    cols_label(
      educ4 = "Education",
      Dem = "DEM.",
      Rep = "REP.",
      Und = "UND."
      ) %>%
  
    fmt_percent(columns = vars(Dem, Rep, Und),
                decimals = 0) %>% 
  
    # This little pipe is that incantation to take this pretty table, turn it
    # into html, and send it to the md file we are creating. Future versions of
    # gt will probably have a better way of doing this. Indeed, does anyone know
    # of one?

    as_raw_html() %>% as.character() %>% cat()

```